A meta analysis of tournaments and an evaluation of performance in the Iterated Prisoner's Dilemma.

### 2019-12-28 paper/main.tex
measure to feature in labels

The result summary, Table~\ref{table:output_result}, has \(N\) rows
because each row contains information for each strategy that participated in the
tournament. The information includes the strategy's rank, median score, the rate
with which the strategy cooperated $(C_r)$, its match win count and the
probability that the strategy cooperated in the opening move. Moreover, the
probabilities of a strategy being in any of the four states ($CC, CD, DC, DD$),
and the rate of which the strategy cooperated after each state. A measure that
has been manually included is the \textbf{normalised rank}. The normalised rank,
denoted as $r$, is calculated as a strategy's rank divided by the tournament's
size ($N$). In the next section the performance of these strategies is evaluated
based on their normalised rank.

@@

The aim of this section is to explore the features that contribute to a
strategy's successful performance. The features explored are measures regarding a
strategy's behaviour, along with measures regarding the tournaments the
strategies competed in. These are given in Table~\ref{table:manual_measures}.

\newcolumntype{g}{>{\columncolor{Gray}}c}
\begin{table}[h]
    \begin{center}
    \resizebox{\textwidth}{!}{
    \begin{tabular}{gcgcgc}
    \toprule
    measure & measure explanation &  source & value type & min value & max value \\
    \midrule
stochastic  &  If a strategy is stochastic & strategy classifier from APL & boolean  & Na &  Na \\
makes use of game &  If a strategy makes used of the game information & strategy classifier from APL & boolean  & Na &  Na \\
makes use of length &  If a strategy makes used of the number of turns & strategy classifier from APL & boolean  & Na &  Na \\
memory usage &  The memory size of a strategy divided by the number of turns & memory size from APL & float & 0 &  1 \\
SSE & A measure of how far a strategy is from ZD behaviour & method described in~\cite{Knight2019} & float & 0 & 1 \\
max cooperating rate $(C_{\text{max}})$  & The biggest cooperating rate in a given tournament  & result summary  & float & 0 & 1\\
min cooperating rate $(C_{\text{min}})$ & The smallest cooperating rate in a given tournament  & result summary  & float & 0 & 1\\
median cooperating rate $(C_{\text{median}})$ & The median cooperating rate in a given tournament  & result summary  & float & 0 & 1\\
mean cooperating rate $(C_{\text{mean}})$ & The mean cooperating rate in a given tournament  & result summary  & float & 0 & 1 \\
$C_r$ / $C_{\text{max}}$ & A strategy's cooperating rate divided by the maximum & result summary  & float & 0 & 1 \\
$C_r$ / $C_{\text{min}}$ & A strategy's cooperating rate divided by the minimum & result summary  & float & 0 & 1 \\
$C_r$ / $C_{\text{median}}$ & A strategy's cooperating rate divided by the median  & result summary  & float & 0 & 1\\
$C_r$ / $C_{\text{mean}}$ & A strategy's cooperating rate divided by the mean & result summary  & float & 0 & 1 \\
$C_r$ & The cooperating ratio of a strategy & result summary  & float & 0 & 1 \\
$CC$ to $C$ rate & The probability a strategy will cooperate after a mutual cooperation & result summary  & float & 0 & 1\\
$CD$ to $C$ rate & The probability a strategy will cooperate after being betrayed by the opponent & result summary  & float & 0 & 1 \\
$DC$ to $C$ rate & The probability a strategy will cooperate after betraying the opponent & result summary  & float & 0 & 1 \\
$DD$ to $C$ rate & The probability a strategy will cooperate after a mutual defection & result summary  & float & 0 & 1 \\
$p_n$ & The probability of a player's action being flip at each interaction & trial summary & float & 0 & 1 \\
$n$ & The number of turns & trial summary & integer & 1 & 200 \\
$p_e$ & The probability of a match ending in the next turn & trial summary & float & 0 & 1 \\
$N$ & The number of strategies in the tournament & trial summary & integer & 3 & 195 \\
$k$ & The number of repetitions of a given tournament & trial summary & integer & 10 & 100 \\
    \bottomrule
        \end{tabular}}
    \end{center}
    \caption{The features which are included in the performance evaluation analysis.}
    \label{table:manual_measures}
\end{table}

APL makes use of classifiers to classify strategies according to
various dimensions. These determine whether
a strategy is stochastic or deterministic, whether it makes use of the number of
turns or the game's payoffs. The memory usage measure is calculated as the
memory size of strategy (which is specified in the strategies implementation
in the APL) divide by the number of turns. For example, Evolved
FSM 16 Noise 05 has a memory size of 16 and participated in a tournament where
$n$ was 134. In the given tournament Evolved FSM 16 Noise 05 has a memory usage
of 0.119. For tournaments with a probabilistic ending the number of turns was
not collected, so the memory usage measure is not used for probabilistic ending
tournaments. The SSE is a measure introduced in~\cite{Knight2019} which shows
how close a strategy is to behaving as a ZDs, and subsequently, in an
extortionate way. The method identifies the ZDs closest to a given strategy and
calculates the algebraic distance between them, defined as SSE. A SSE value of 1
indicates no extortionate behaviour at all whereas a value of 0 indicates that a
strategy is behaving a ZDs. The rest of the features considered are the $CC$ to
$C$, $CD$ to $C$, $DC$ to $C$, and $DD$ to $C$ rates as well as cooperating
ratio of a strategy. The minimum, maximum, medium and median cooperating ratios
of each tournament are also included, and finally the number of turns, the
number of strategies, the number of repetitions and the probabilities of noise
and the game ending are also included.

Table~\ref{table:correlations} shows the correlation coefficients between the
features of Table~\ref{table:manual_measures} the median score and the median
normalised rank. Note that the correlation for the classifiers is
not included because they are binary variables and they will be evaluated using a
different method. The correlation coefficients for all the features in
Table~\ref{table:manual_measures} against themselves have also been calculated and a
graphical representation can be found in the Appendix~\ref{app:correlations}.

\newcolumntype{g}{>{\columncolor{Gray}}c}
\begin{table}[!htbp]
    \begin{center}
    \resizebox{.9\textwidth}{!}{
        \input{correlations_table.tex}
    }
\end{center}
\caption{Correlations table between the features of Table~\ref{table:manual_measures}
the normalised rank and the median score.}\label{table:correlations}
\end{table}

@@

The importance that the features of Table~\ref{table:manual_measures} had on
each classification task; to which cluster a performance was assigned to based
on the normalised rank, and their normalised rank and median score have been
calculated and are given by Figures~\ref{fig:clustering_importance_standard},
\ref{fig:clustering_importance_noise},~\ref{fig:clustering_importance_probend},
\ref{fig:clustering_importance_probend_noise}
and~\ref{fig:clustering_importance_overall}. These show that the classifiers
stochastic, make use of game and make use of length have no significant effect,
and several of the features that are highted by the importance are inline with
the correlation results. Moreover, the smoothing parameter \(k\) appears to no
have a significant effect either. The most important features based on the
random forest analysis were $C_{r} / C_{median}$ and $C_r / C_{mean}$.

@@

\begin{table}[h]
    \begin{center}
    \resizebox{.7\textwidth}{!}{
    \begin{tabular}{llc}
    \toprule
measure & measure explanation \\
 \midrule
stochastic                  & If a strategy is stochastic \\
makes use of game           & If a strategy makes used of the game information  \\
makes use of length         & If a strategy makes used of the number of turns \\
memory usage                & The memory size of a strategy divided by the number of turns \\
SSE                         & A measure of how far a strategy is from extortionate behaviour \\
$C_{\text{max}}$          & The biggest cooperating rate in the tournament \\
$C_{\text{min}}$          & The smallest cooperating rate in the tournament \\
$C_{\text{median}}$       & The median cooperating rate in the tournament \\
$C_{\text{mean}}$         & The mean cooperating rate in the tournament \\
$C_r$ / $C_{\text{max}}$    & A strategy's cooperating rate divided by the maximum \\
$C_r$ / $C_{\text{min}}$    & A strategy's cooperating rate divided by the minimum \\
$C_r$ / $C_{\text{median}}$ & A strategy's cooperating rate divided by the median \\
$C_r$ / $C_{\text{mean}}$   & A strategy's cooperating rate divided by the mean \\
$C_r$                       & The cooperating ratio of a strategy \\
$CC$ to $C$ rate            & The probability a strategy will cooperate after a mutual cooperation \\
$CD$ to $C$ rate            & The probability a strategy will cooperate after being betrayed by the opponent \\
$DC$ to $C$ rate            & The probability a strategy will cooperate after betraying the opponent\\
$DD$ to $C$ rate            & The probability a strategy will cooperate after a mutual defection\\
$p_n$                       & The probability of a player's action being flip at each interaction \\
$n$                         & The number of turns \\
$p_e$                       & The probability of a match ending in the next turn \\
$N$                         & The number of strategies in the tournament \\
$k$                         & The number that a given tournament is repeated \\
    \bottomrule
        \end{tabular}}
    \end{center}
    \caption{The features which are included in the performance evaluation analysis.}\label{table:parameters_summary}
\end{table}

@@

A graphical representation of the correlation coefficients for the features in
Table~\ref{table:manual_measures}.

\begin{figure}[!htbp]
        \begin{center}
            \includegraphics[width=.75\linewidth]{../images/standard_correlation_plot.pdf}
        \end{center}
        \caption{Correlation coefficients of features in Table~\ref{table:manual_measures}
        for standard tournaments}
\end{figure}
\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.75\linewidth]{../images/noise_correlation_plot.pdf}
    \end{center}
    \caption{Correlation coefficients of features in Table~\ref{table:manual_measures}
    for noisy tournaments}
\end{figure}
\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.75\linewidth]{../images/probend_correlation_plot.pdf}
    \end{center}
    \caption{Correlation coefficients of features in Table~\ref{table:manual_measures}
    for probabilistic ending tournaments}
\end{figure}
\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.75\linewidth]{../images/probend_noise_correlation_plot.pdf}
    \end{center}
    \caption{Correlation coefficients of features in Table~\ref{table:manual_measures}
    for noisy probabilistic ending tournaments}
\end{figure}
\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.75\linewidth]{../images/merged_correlation_plot.pdf}
    \end{center}
    \caption{Correlation coefficients of features in Table~\ref{table:manual_measures}
    for data set}
\end{figure}

>>>>

The result summary, Table~\ref{table:output_result}, has \(N\) rows
because each row contains information for each strategy that participated in the
tournament. The information includes the strategy's rank, median score, the rate
with which the strategy cooperated $(C_r)$, its match win count and the
probability that the strategy cooperated in the opening move. Moreover, the
probabilities of a strategy being in any of the four states ($CC, CD, DC, DD$),
and the rate of which the strategy cooperated after each state. A feature that
has been manually included is the \textbf{normalised rank}. The normalised rank,
denoted as $r$, is calculated as a strategy's rank divided by the tournament's
size ($N$). In the next section the performance of these strategies is evaluated
based on their normalised rank.

@@

The aim of this section is to explore the features that contribute to a
strategy's successful performance. The features explored are measures regarding a
strategy's behaviour, along with measures regarding the tournaments the
strategies competed in. These are given in Table~\ref{table:manual_features}.

\newcolumntype{g}{>{\columncolor{Gray}}c}
\begin{table}[h]
    \begin{center}
    \resizebox{\textwidth}{!}{
    \begin{tabular}{gcgcgc}
    \toprule
    feature & feature explanation &  source & value type & min value & max value \\
    \midrule
stochastic  &  If a strategy is stochastic & strategy classifier from APL & boolean  & Na &  Na \\
makes use of game &  If a strategy makes used of the game information & strategy classifier from APL & boolean  & Na &  Na \\
makes use of length &  If a strategy makes used of the number of turns & strategy classifier from APL & boolean  & Na &  Na \\
memory usage &  The memory size of a strategy divided by the number of turns & memory size from APL & float & 0 &  1 \\
SSE & A measure of how far a strategy is from ZD behaviour & method described in~\cite{Knight2019} & float & 0 & 1 \\
max cooperating rate $(C_{\text{max}})$  & The biggest cooperating rate in a given tournament  & result summary  & float & 0 & 1\\
min cooperating rate $(C_{\text{min}})$ & The smallest cooperating rate in a given tournament  & result summary  & float & 0 & 1\\
median cooperating rate $(C_{\text{median}})$ & The median cooperating rate in a given tournament  & result summary  & float & 0 & 1\\
mean cooperating rate $(C_{\text{mean}})$ & The mean cooperating rate in a given tournament  & result summary  & float & 0 & 1 \\
$C_r$ / $C_{\text{max}}$ & A strategy's cooperating rate divided by the maximum & result summary  & float & 0 & 1 \\
$C_r$ / $C_{\text{min}}$ & A strategy's cooperating rate divided by the minimum & result summary  & float & 0 & 1 \\
$C_r$ / $C_{\text{median}}$ & A strategy's cooperating rate divided by the median  & result summary  & float & 0 & 1\\
$C_r$ / $C_{\text{mean}}$ & A strategy's cooperating rate divided by the mean & result summary  & float & 0 & 1 \\
$C_r$ & The cooperating ratio of a strategy & result summary  & float & 0 & 1 \\
$CC$ to $C$ rate & The probability a strategy will cooperate after a mutual cooperation & result summary  & float & 0 & 1\\
$CD$ to $C$ rate & The probability a strategy will cooperate after being betrayed by the opponent & result summary  & float & 0 & 1 \\
$DC$ to $C$ rate & The probability a strategy will cooperate after betraying the opponent & result summary  & float & 0 & 1 \\
$DD$ to $C$ rate & The probability a strategy will cooperate after a mutual defection & result summary  & float & 0 & 1 \\
$p_n$ & The probability of a player's action being flip at each interaction & trial summary & float & 0 & 1 \\
$n$ & The number of turns & trial summary & integer & 1 & 200 \\
$p_e$ & The probability of a match ending in the next turn & trial summary & float & 0 & 1 \\
$N$ & The number of strategies in the tournament & trial summary & integer & 3 & 195 \\
$k$ & The number of repetitions of a given tournament & trial summary & integer & 10 & 100 \\
    \bottomrule
        \end{tabular}}
    \end{center}
    \caption{The features which are included in the performance evaluation analysis.}
    \label{table:manual_features}
\end{table}

APL makes use of classifiers to classify strategies according to
various dimensions. These determine whether
a strategy is stochastic or deterministic, whether it makes use of the number of
turns or the game's payoffs. The memory usage feature is calculated as the
memory size of strategy (which is specified in the strategies implementation
in the APL) divide by the number of turns. For example, Evolved
FSM 16 Noise 05 has a memory size of 16 and participated in a tournament where
$n$ was 134. In the given tournament Evolved FSM 16 Noise 05 has a memory usage
of 0.119. For tournaments with a probabilistic ending the number of turns was
not collected, so the memory usage feature is not used for probabilistic ending
tournaments. The SSE is a feature introduced in~\cite{Knight2019} which shows
how close a strategy is to behaving as a ZDs, and subsequently, in an
extortionate way. The method identifies the ZDs closest to a given strategy and
calculates the algebraic distance between them, defined as SSE. A SSE value of 1
indicates no extortionate behaviour at all whereas a value of 0 indicates that a
strategy is behaving a ZDs. The rest of the features considered are the $CC$ to
$C$, $CD$ to $C$, $DC$ to $C$, and $DD$ to $C$ rates as well as cooperating
ratio of a strategy. The minimum, maximum, medium and median cooperating ratios
of each tournament are also included, and finally the number of turns, the
number of strategies, the number of repetitions and the probabilities of noise
and the game ending are also included.

Table~\ref{table:correlations} shows the correlation coefficients between the
features of Table~\ref{table:manual_features} the median score and the median
normalised rank. Note that the correlation for the classifiers is
not included because they are binary variables and they will be evaluated using a
different method. The correlation coefficients for all the features in
Table~\ref{table:manual_features} against themselves have also been calculated and a
graphical representation can be found in the Appendix~\ref{app:correlations}.

\newcolumntype{g}{>{\columncolor{Gray}}c}
\begin{table}[!htbp]
    \begin{center}
    \resizebox{.9\textwidth}{!}{
        \input{correlations_table.tex}
    }
\end{center}
\caption{Correlations table between the features of Table~\ref{table:manual_features}
the normalised rank and the median score.}\label{table:correlations}
\end{table}

@@

The importance that the features of Table~\ref{table:manual_features} had on
each classification task; to which cluster a performance was assigned to based
on the normalised rank, and their normalised rank and median score have been
calculated and are given by Figures~\ref{fig:clustering_importance_standard},
\ref{fig:clustering_importance_noise},~\ref{fig:clustering_importance_probend},
\ref{fig:clustering_importance_probend_noise}
and~\ref{fig:clustering_importance_overall}. These show that the classifiers
stochastic, make use of game and make use of length have no significant effect,
and several of the features that are highted by the importance are inline with
the correlation results. Moreover, the smoothing parameter \(k\) appears to no
have a significant effect either. The most important features based on the
random forest analysis were $C_{r} / C_{median}$ and $C_r / C_{mean}$.

@@

\begin{table}[h]
    \begin{center}
    \resizebox{.7\textwidth}{!}{
    \begin{tabular}{llc}
    \toprule
feature & feature explanation \\
 \midrule
stochastic                  & If a strategy is stochastic \\
makes use of game           & If a strategy makes used of the game information  \\
makes use of length         & If a strategy makes used of the number of turns \\
memory usage                & The memory size of a strategy divided by the number of turns \\
SSE                         & A measure of how far a strategy is from extortionate behaviour \\
$C_{\text{max}}$          & The biggest cooperating rate in the tournament \\
$C_{\text{min}}$          & The smallest cooperating rate in the tournament \\
$C_{\text{median}}$       & The median cooperating rate in the tournament \\
$C_{\text{mean}}$         & The mean cooperating rate in the tournament \\
$C_r$ / $C_{\text{max}}$    & A strategy's cooperating rate divided by the maximum \\
$C_r$ / $C_{\text{min}}$    & A strategy's cooperating rate divided by the minimum \\
$C_r$ / $C_{\text{median}}$ & A strategy's cooperating rate divided by the median \\
$C_r$ / $C_{\text{mean}}$   & A strategy's cooperating rate divided by the mean \\
$C_r$                       & The cooperating ratio of a strategy \\
$CC$ to $C$ rate            & The probability a strategy will cooperate after a mutual cooperation \\
$CD$ to $C$ rate            & The probability a strategy will cooperate after being betrayed by the opponent \\
$DC$ to $C$ rate            & The probability a strategy will cooperate after betraying the opponent\\
$DD$ to $C$ rate            & The probability a strategy will cooperate after a mutual defection\\
$p_n$                       & The probability of a player's action being flip at each interaction \\
$n$                         & The number of turns \\
$p_e$                       & The probability of a match ending in the next turn \\
$N$                         & The number of strategies in the tournament \\
$k$                         & The number that a given tournament is repeated \\
    \bottomrule
        \end{tabular}}
    \end{center}
    \caption{The features which are included in the performance evaluation analysis.}\label{table:parameters_summary}
\end{table}

@@

A graphical representation of the correlation coefficients for the features in
Table~\ref{table:manual_features}.

\begin{figure}[!htbp]
        \begin{center}
            \includegraphics[width=.75\linewidth]{../images/standard_correlation_plot.pdf}
        \end{center}
        \caption{Correlation coefficients of features in Table~\ref{table:manual_features}
        for standard tournaments}
\end{figure}
\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.75\linewidth]{../images/noise_correlation_plot.pdf}
    \end{center}
    \caption{Correlation coefficients of features in Table~\ref{table:manual_features}
    for noisy tournaments}
\end{figure}
\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.75\linewidth]{../images/probend_correlation_plot.pdf}
    \end{center}
    \caption{Correlation coefficients of features in Table~\ref{table:manual_features}
    for probabilistic ending tournaments}
\end{figure}
\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.75\linewidth]{../images/probend_noise_correlation_plot.pdf}
    \end{center}
    \caption{Correlation coefficients of features in Table~\ref{table:manual_features}
    for noisy probabilistic ending tournaments}
\end{figure}
\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.75\linewidth]{../images/merged_correlation_plot.pdf}
    \end{center}
    \caption{Correlation coefficients of features in Table~\ref{table:manual_features}
    for data set}
\end{figure}
